---
title: "DM Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(RSQLite)
library(DBI)
library(lubridate)

# Load the scales package for formatting scales and axes
library(scales)
```

## Create and Connect Database
```{r}
# Define the tables to drop
tables_to_drop <- c("ads", "category", "customer_address", "customer", 
                     "delivery", "orders", "orders_details", "product", 
                     "supplier", "supply", "`transaction`")

# Connect to the database
my_connection <- RSQLite::dbConnect(RSQLite::SQLite(), "e_commerce.db")

# Drop each table
for (table_name in tables_to_drop) {
  dbExecute(my_connection, paste("DROP TABLE IF EXISTS", table_name))
}

# After dropping tables, you can proceed to create and load new tables
sql_commands <- readLines("Data analysis/Schema.sql", warn = FALSE)
for (sql_cmd in sql_commands) {
  tryCatch({
    dbExecute(my_connection, sql_cmd)
  }, error = function(e) {
    # Print error message without stopping the execution
    #cat("Error occurred: ", conditionMessage(e), "\n")
  })
}
```

## Import Data
```{r}
ads = readr::read_csv('ecommerce_data/ads_data.csv', col_types = cols(ad_start_date = "c", ad_end_date = "c"))
category = readr::read_csv('ecommerce_data/category_data.csv')
customer_address = readr::read_csv('ecommerce_data/customer_addresses_data.csv')
customer = readr::read_csv('ecommerce_data/customers_data.csv', col_types = cols(date_of_birth = "c"))
delivery = readr::read_csv('ecommerce_data/delivery_data.csv', col_types = cols(delivery_start_date = "c", delivery_end_date = "c"))
orders_details = readr::read_csv('ecommerce_data/order_details_data.csv')
orders = readr::read_csv('ecommerce_data/orders_data.csv', col_types = cols(order_date = "c"))
product = readr::read_csv('ecommerce_data/product_data.csv')
supplier = readr::read_csv('ecommerce_data/supplier_data.csv')
supply = readr::read_csv('ecommerce_data/supply_data.csv')
transaction = readr::read_csv('ecommerce_data/transactions_data.csv', col_types = cols(transaction_time = "c"))
```

## Data Quality Check before writing into database
```{r}
# Ads
str(ads) 
ads$ad_status = as.factor(ads$ad_status)
str(ads)
summary(ads)

# Category
# Missing value in parent_category_id is reasonable if category_id is parent_category_id itself
str(category)
summary(category)

# Customer_address
str(customer_address)
customer_address$country = as.factor(customer_address$country)
customer_address$state = as.factor(customer_address$state)
customer_address$city = as.factor(customer_address$city)
str(customer_address)
summary(customer_address)

# Customer
str(customer)
customer$gender = as.factor(customer$gender)
str(customer)
summary(customer)

# Delivery
# It's reasonable to have missing values for delivery_start_date and delivery_end_date in "Failed" and "Not_Delivered" delivery_status
str(delivery)
delivery$delivery_status = as.factor(delivery$delivery_status)
str(delivery)
summary(delivery)

# orders
str(orders)
orders$order_status = as.factor(orders$order_status)
str(orders)
summary(orders)

# orders_Details
str(orders_details)
summary(orders_details)

# Product
str(product)
summary(product)

# Supplier
str(supplier)
supplier$supplier_phone = as.character(supplier$supplier_phone)
supplier$supplier_country = as.factor(supplier$supplier_country)
supplier$supplier_state = as.factor(supplier$supplier_state)
supplier$supplier_city = as.factor(supplier$supplier_city)
str(supplier)
summary(supplier)

# Supply
str(supply)
summary(supply)

# Transaction
str(transaction)
# Transform transaction_time without the float
transaction$transaction_time = sub("\\..*", "", transaction$transaction_time)
transaction$payment_method = as_factor(transaction$payment_method)
transaction$transaction_status = as_factor(transaction$transaction_status)
str(transaction)
summary(transaction)
```

## Write them to the database 
```{r}
RSQLite::dbWriteTable(my_connection,"ads",ads,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"category",category,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"customer_address",customer_address,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"customer",customer,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"delivery",delivery,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"orders",orders,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"orders_details",orders_details,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"product",product,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"supplier",supplier,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"supply",supply,append = TRUE,overwrite=FALSE)
RSQLite::dbWriteTable(my_connection,"transaction",transaction,append = TRUE,overwrite=FALSE)
```

## Database check
```{r}
# Tables check to see if there's any error in the data import stage
ads_check = tbl(my_connection, "ads")
ads_check
category_check = tbl(my_connection, "category")
category_check
customer_address_check = tbl(my_connection, "customer_address")
customer_address_check
customer_check = tbl(my_connection, "customer")
customer_check
delivery_check = tbl(my_connection, "delivery")
delivery_check
orders_check = tbl(my_connection, "orders")
orders_check
orders_details_check = tbl(my_connection, "orders_details")
orders_details_check
product_check = tbl(my_connection, "product")
product_check
supplier_check = tbl(my_connection, "supplier")
supplier_check
supply_check = tbl(my_connection, "supply")
supply_check
transaction_check = tbl(my_connection, "transaction")
transaction_check
```
It seems no data structure errors in this stage.

---------------------------------------------------------------------------

# Analysis

## 1 Product Popularity 
```{r}
# Execute SQL query to count the frequency of each product_id and retrieve top 10 popular products
top_10_popular_products <- dbGetQuery(my_connection, "
  SELECT p.product_name, 
         SUM(od.sub_quantity) AS Frequency
  FROM orders_details od
  JOIN product p ON od.product_id = p.product_id
  GROUP BY od.product_id, p.product_name
  ORDER BY Frequency DESC
  LIMIT 10
")

# Plot the bar chart of the top 10 popular products with product names on the x-axis
ggplot(top_10_popular_products, aes(x = reorder(product_name, -Frequency), y = Frequency)) +
  geom_bar(stat = "identity", fill = "#6495ED") +
  geom_text(aes(label = Frequency), vjust = -0.5, size = 3, color = "black") +  # Add numeric labels on top of bars
  labs(title = "Top 10 Popular Products",
       x = "Product Name",
       y = "Amount Consumed") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2 Product Sales Analysis 
```{r}
# Execute SQL query to calculate total revenue for each product and retrieve top 10 products
product_total_revenue <- dbGetQuery(my_connection, "
  SELECT product.product_id, 
         product.product_name,
         SUM(CASE 
               WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
               ELSE orders_details.sub_quantity * product.product_price
             END) AS total_revenue
  FROM orders_details
  JOIN product ON orders_details.product_id = product.product_id
  JOIN orders ON orders_details.order_id = orders.order_id
  JOIN customer ON orders.customer_id = customer.customer_id
  GROUP BY product.product_id
  ORDER BY total_revenue DESC
  LIMIT 10
")

# Create the ggplot for total revenue with product names on the x-axis
ggplot(product_total_revenue, aes(x = reorder(product_name, desc(total_revenue)), y = total_revenue)) +
  geom_bar(stat = "identity", fill = "#6495ED") +
  labs(title = "Total Sales Figures for Top 10 Products",
       x = "Product Name",
       y = "Total Sales") +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(product_total_revenue$total_revenue), by = 25000)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 3 Category Sales & Popularity Analysis
```{r}
# Category Popularity
category_popularity <- dbGetQuery(my_connection, "
SELECT SUBSTRING(product.category_id, 1, 6) AS category,
       SUM(orders_details.sub_quantity) AS total_quantity,
       category.category_name
FROM product
JOIN orders_details ON product.product_id = orders_details.product_id
JOIN category ON SUBSTRING(product.category_id, 1, 6) = category.category_id
GROUP BY SUBSTRING(product.category_id, 1, 6), category.category_name
")

# Category Sales
category_sales <- dbGetQuery(my_connection, "
SELECT SUBSTRING(product.category_id, 1, 6) AS category,
       SUM(CASE
               WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
               ELSE orders_details.sub_quantity * product.product_price
           END) AS total_sales,
       category.category_name
FROM product
JOIN orders_details ON product.product_id = orders_details.product_id
JOIN orders ON orders_details.order_id = orders.order_id
JOIN customer ON orders.customer_id = customer.customer_id
JOIN category ON SUBSTRING(product.category_id, 1, 6) = category.category_id
GROUP BY SUBSTRING(product.category_id, 1, 6), category.category_name
")

# Combine category sales and category popularity data, removing duplicate category_name column
combined_data <- inner_join(category_sales, category_popularity %>% select(-category_name), by = "category")

# Create the plot
ggplot(combined_data, aes(x = category_name)) +
  geom_bar(aes(y = total_sales), stat = "identity", fill = "#6495ED") +  # Bar plot representing category sales
  geom_line(aes(y = total_quantity * 1000), color = "red", group = 1) +  # Line plot representing category popularity, multiplied by 1000 just for demonstration of comparison on the same plot
  scale_y_continuous(name = "Total Sales", sec.axis = sec_axis(~./1000, name = "Popularity")) +  # Set dual y-axes, representing sales and popularity
  labs(title = "Category Sales and Popularity Analysis",
       x = "Category", y = "Total Sales",) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 4 Gender & Category Visualization
```{r}
# Execute SQL query in R
category_gender_sales <- dbGetQuery(my_connection, "
SELECT SUBSTRING(product.category_id, 1, 6) AS category,
       customer.gender,
       SUM(CASE
               WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
               ELSE orders_details.sub_quantity * product.product_price
           END) AS total_sales,
       category.category_name
FROM product
JOIN orders_details ON product.product_id = orders_details.product_id
JOIN orders ON orders_details.order_id = orders.order_id
JOIN customer ON orders.customer_id = customer.customer_id
JOIN category ON SUBSTRING(product.category_id, 1, 6) = category.category_id
GROUP BY SUBSTRING(product.category_id, 1, 6), category.category_name, customer.gender
")

# Plot gender-category sales relationship
ggplot(category_gender_sales, aes(x = category_name, y = total_sales, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Gender & Category Sales Analysis",
       x = "Category Name",
       y = "Total Sales",
       fill = "Gender") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## 5 Age Visualization
```{r}
customer1 <- dbGetQuery(my_connection, "
  SELECT customer.*,
  FLOOR((julianday('now') - julianday(customer.date_of_birth)) / 365.25) AS age1
  FROM customer")

# Plot the distribution of Age
ggplot(customer1, aes(x = age1)) +
  geom_histogram(aes(y = after_stat(density)), fill = "#6495ED", binwidth = 5) +
  geom_density(color = "red") +
  labs(title = "Histogram and Density Plot of Age", x = "Age", y = "Density") +
  scale_x_continuous(breaks = seq(0, max(customer1$age1), by = 5)) +
  theme_minimal()
```

## 6 Age & Category Visualization (pass)
```{r}
# Execute SQL query to get relevant data
customer_data <- dbGetQuery(my_connection, "
  SELECT FLOOR((julianday('now') - julianday(customer.date_of_birth)) / 365.25) AS age,
         SUM(CASE
                 WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
                 ELSE orders_details.sub_quantity * product.product_price
             END) AS total_sales,
         SUBSTRING(product.category_id, 1, 6) AS category,
         category.category_name
  FROM customer
  JOIN orders ON customer.customer_id = orders.customer_id
  JOIN orders_details ON orders.order_id = orders_details.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN category ON SUBSTRING(product.category_id, 1, 6) = category.category_id
  GROUP BY age, customer.gender, category, category.category_name
")

# Define age groups with intervals of 10 years
customer_data$age_group <- cut(customer_data$age, breaks = seq(0, max(customer_data$age) + 10, by = 10))

# Plot Age & Category Visualization
ggplot(customer_data, aes(x = age_group, y = total_sales, fill = category_name)) +
  geom_bar(position = "stack", stat = "identity") +
  labs(title = "Total Sales by Age Group and Category",
       x = "Age Group",
       y = "Total Sales",
       fill = "Category") +
  theme_minimal()
```

## 7 Region Revenue Analysis
```{r}
# Execute SQL query to calculate total revenue for each state
state_total_revenue <- dbGetQuery(my_connection, "
  SELECT customer_address.state, 
         SUM(
           CASE 
             WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
             ELSE orders_details.sub_quantity * product.product_price
           END
         ) AS total_revenue
  FROM orders
  JOIN orders_details ON orders.order_id = orders_details.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN customer ON orders.customer_id = customer.customer_id
  JOIN customer_address ON customer.customer_id = customer_address.customer_id
  GROUP BY customer_address.state
")

# Select the top 10 states by total revenue
top_10_state_total_revenue <- state_total_revenue %>%
  arrange(desc(total_revenue)) %>%
  head(10)

# Plot the total revenue for the top 10 states
ggplot(top_10_state_total_revenue, aes(x = reorder(state, -total_revenue), y = total_revenue)) + 
  geom_bar(stat = "identity", fill = "#6495ED") +
  labs(title = "Top 10 States by Total Consumption",
       x = "State",
       y = "Total Consumption") +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma, breaks = seq(0, max(top_10_state_total_revenue$total_revenue), by = 250000)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

--------------------------------------------------------------------

# Time Analysis

## 8 Total Order Value Trend Analysis
```{r}
# Convert Order_Date to month-year format directly in SQL and calculate total order value per month
total_order_value_by_month <- dbGetQuery(my_connection, "
  SELECT strftime('%Y-%m', orders.order_date) AS month_year,
         SUM(
           CASE 
             WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
             ELSE orders_details.sub_quantity * product.product_price
           END
         ) AS total_order_value
  FROM orders
  JOIN orders_details ON orders.order_id = orders_details.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN customer ON orders.customer_id = customer.customer_id
  GROUP BY month_year
")

# Plot the trend of total order value over time
ggplot(total_order_value_by_month, aes(x = as.Date(paste0(month_year, "-01"), "%Y-%m-%d"), y = total_order_value, group = 1)) +
  geom_line(color = "#6495ED", size = 1, aes(group = 1)) +
  labs(title = "Trend of Total Order Value Over Time",
       x = "Month",
       y = "Total Order Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_y_continuous(labels = scales::dollar) +  # Adjust y-axis labels to use dollar formatting
  scale_x_date(labels = scales::date_format("%b %Y")) +  # Format x-axis labels as Sep 2023
  theme(legend.position = "none")  # Remove legend to avoid redundancy
```

## 9 Average Order Value Trend Analysis 
```{r}
# Convert Order_Date to month-year format directly in SQL and calculate average order value per month
average_order_value_by_month <- dbGetQuery(my_connection, "
  SELECT strftime('%Y-%m', orders.order_date) AS month_year,
         AVG(
           CASE 
             WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
             ELSE orders_details.sub_quantity * product.product_price
           END
         ) AS average_order_value
  FROM orders
  JOIN orders_details ON orders.order_id = orders_details.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN customer ON orders.customer_id = customer.customer_id
  GROUP BY month_year
")

# Plot the trend of average order value over time
ggplot(average_order_value_by_month, aes(x = as.Date(paste0(month_year, "-01"), "%Y-%m-%d"), y = average_order_value, group = 1)) +
  geom_line(color = "#6495ED", size = 1, aes(group = 1)) +
  labs(title = "Trend of Average Order Value Over Time",
       x = "Month",
       y = "Average Order Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_y_continuous(labels = scales::dollar) +  # Adjust y-axis labels to use dollar formatting
  scale_x_date(labels = scales::date_format("%b %Y")) +  # Format x-axis labels as Sep 2023
  theme(legend.position = "none")  # Remove legend to avoid redundancy

```

## 10 Category Sales Trend Analysis 
```{r}
# Execute SQL query in R
category_sales_time <- dbGetQuery(my_connection, "
SELECT strftime('%Y-%m', orders.order_date) AS category_month_year,
       SUBSTRING(product.category_id, 1, 6) AS category,
       SUM(CASE
               WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
               ELSE orders_details.sub_quantity * product.product_price
           END) AS total_sales,
       category.category_name
FROM product
JOIN orders_details ON product.product_id = orders_details.product_id
JOIN orders ON orders_details.order_id = orders.order_id
JOIN customer ON orders.customer_id = customer.customer_id
JOIN category ON SUBSTRING(product.category_id, 1, 6) = category.category_id
GROUP BY strftime('%Y-%m', orders.order_date), SUBSTRING(product.category_id, 1, 6), category.category_name
")

# Convert category_month_year to Date format
category_sales_time$category_month_year <- as.Date(paste0(category_sales_time$category_month_year, "-01"))

# Plot the line chart of category sales over time for the five main categories
ggplot(category_sales_time, aes(x = category_month_year, y = total_sales, color = category_name)) + geom_line() +
  labs(title = "Category Sales Analysis Over Time",
       x = "Month",
       y = "Total Sales",
       color = "Category Name") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::comma)
```

## 11 Membership Revenue Trend Analysis
```{r}
# Execute SQL query in the database and fetch the results
membership_revenue_by_month <- dbGetQuery(my_connection, "
  SELECT strftime('%Y-%m', orders.order_date) AS membership_month_year,
         customer.membership,
         SUM(CASE 
               WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
               ELSE orders_details.sub_quantity * product.product_price
             END) AS total_revenue
  FROM orders
  JOIN orders_details ON orders.order_id = orders_details.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN customer ON orders.customer_id = customer.customer_id
  GROUP BY membership_month_year, customer.membership
")

# Convert the date column to a date-time format
membership_revenue_by_month$membership_month_year <- ym(membership_revenue_by_month$membership_month_year)

# Plot the trend of membership revenue over time
ggplot(membership_revenue_by_month, aes(x = membership_month_year, y = total_revenue, group = membership, color = membership)) +
  geom_line(size = 1) +
  labs(title = "Trend of Membership Consumption Over Time",
       x = "Month",
       y = "Total Consumption",
       color = "Membership Status") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::comma)
```

## 12 Total Product Revenue Analysis by Rate 
```{r}
# Query to retrieve product revenue by rate category
product_revenue_by_rate <- dbGetQuery(my_connection, "
  SELECT strftime('%Y-%m', orders.order_date) AS product_month_year,
         SUM(
           CASE 
             WHEN customer.membership = 'yes' THEN orders_details.sub_quantity * product.product_price * product.discount_percentage
             ELSE orders_details.sub_quantity * product.product_price
           END
         ) AS total_revenue,
         CASE
           WHEN product.rate_value BETWEEN 1 AND 2 THEN '1-2'
           WHEN product.rate_value BETWEEN 2 AND 3 THEN '2-3'
           WHEN product.rate_value BETWEEN 3 AND 4 THEN '3-4'
           WHEN product.rate_value BETWEEN 4 AND 5 THEN '4-5'
           ELSE 'Unknown'
         END AS rate_category
  FROM orders_details
  JOIN orders ON orders_details.order_id = orders.order_id
  JOIN product ON orders_details.product_id = product.product_id
  JOIN customer ON orders.customer_id = customer.customer_id
  GROUP BY product_month_year, rate_category
")

# Convert the date column to a date-time format
product_revenue_by_rate$product_month_year <- ym(product_revenue_by_rate$product_month_year)

# Plot the trend of product revenue by rate category over time
ggplot(product_revenue_by_rate, aes(x = product_month_year, y = total_revenue, group = rate_category, color = rate_category)) +
  geom_line(size = 1) +
  labs(title = "Product Sales Trend by Rate Category Over Time",
       x = "Month",
       y = "Total Product Sales",
       color = "Rate Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::comma)
```


## Disconnect from the database
```{r}
dbDisconnect(my_connection)
```

